A Compound Decision Approach to Covariance Matrix Estimation

### 2021-08-24 msg.tex
changes of ref2 and part of ref1

\title[Empirical Bayes Covariance Matrix Estimation]{A Nonparametric Empirical Bayes Approach to Covariance Matrix Estimation}

@@

Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
\begin{equation}
\label{linear model}
\hat{\bs{\Sigma}}_{\text{LW}} = \hat{\rho}_{\text{LW}} \bs{S} + (1-\hat{\rho}_{\text{LW}})\hat{\mu} \bs{I}
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, and $\hat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\hat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$. Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \ref{frobenius risk}. We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
\begin{equation}
  \label{linear class}
  \beta_S s_{jk} + \beta_I u_{jk},
\end{equation}
where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. It is straightforward to show that
$$\hat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
is an unbiased estimate of the risk \ref{frobenius risk}, where $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$. The optimal estimator in this class can now be chosen by minimizing $\hat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}. Proposition \ref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \ref{linear model}.
\begin{prop}
  \label{prop:linear}
  Define the estimator $\hat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\hat{\sigma}_{jk} = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, where $(\hat{\beta}_S, \hat{\beta}_I) = \argmin_{\beta_S, \beta_I} \hat{R}(\beta_S, \beta_I)$. Then $\hat{\bs{\Sigma}}_V$ = $\hat{\bs{\Sigma}}_{LW}$.
\end{prop}

@@

The previous section argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, but discusses only estimators linear in $s_{jk}$. We propose to consider a larger class, the class of so-called separable rules. In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ \citep{robbins1951asymptotically}. Here we generalize this to the problem of estimating a vectorized matrix. For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, define the class of separable rules
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})\},
\end{equation}
where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature. In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$.

@@

The optimal $\bs{\delta}^\star$ is an oracle estimator and cannot be calculated in practice, as the true risk is unknown. In the classical compound decision framework, empirical Bayes methods are used to estimate the oracle optimal separable rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}. We take a similar approach here. To simplify notation, denote $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})^\top$ as $\bs{A}_{jk}$. The density of $f(\cdot \mid \bs{\eta}_{jk})$ of $\bs{A}_{jk}$ depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation. When $r_{jk} \ne 1$, $\bs{A}_{jk}$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
\[
\bs{C}_{jk}=\begin{bmatrix}
&\sigma_j^2 & \sigma_j\sigma_k r_{jk} \\
& \sigma_j\sigma_k r_{jk}  & \sigma_k^2
\end{bmatrix}.
\]
When $r_{jk} = 1$, $\bs{A}_{jk}$ consists of mean-zero univariate normals with variances $\sigma_j^2$.

Now consider the Bayesian model
\begin{equation}
  \label{bayesian}
  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \ref{frobenius risk}: for any separable $\bs{\delta} \in \mathcal{S}$ \ref{separable}, the Frobenius risk can be written as
\begin{align*}
  R(\bs{\Sigma}, \bs{\delta})
  =\,&
  \frac{1}{p^2} \sum_{j,k = 1}^p
  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A}\\
  =\,&
  \int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_p(\bs{\eta}) d\bs{A}
  =
  \bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2],
\end{align*}
where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$. The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ minimizes the Bayes risk.

Based on this result, we propose the following empirical Bayes procedure. We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the prior $G_p$. Under the Bayesian model \ref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_p$ using
\begin{equation}
  \label{Gp hat}
  \hat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
\end{equation}
where $\mathcal{G}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$. Of course, the $\bs{A}_{jk}$ are not independent, so $\hat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\hat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using
\begin{equation}
  \label{proposed}
  \hat{\bs{\delta}}(\bs{X})
  =
  (\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
  \quad
  \hat{t}(\bs{A}) = \frac{\int g(\bs{\eta}) f(\bs{A} \mid \bs{\eta})d\hat{G}_p(\bs{\eta})}{\int f(\bs{A} \mid \bs{\eta}) d\hat{G}_p(\bs{\eta})}.
\end{equation}
The $\hat{t}$ estimates the Bayes rule $t^\star$ and $\hat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.

Our proposed procedure is an example of what \citet{efron2014two} calls $g$-modeling, an approach to empirical Bayes problems that proceeds by modeling the prior. A major advantage of nonparametric estiation of the prior is that it allows the data itself to determine how best to shrink the estimator. In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}. Theoretical justification of our proposed $\hat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}. Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\hat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.

@@

Calculating the estimated prior $\hat{G}_p$ \ref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$. \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points. The EM algorithm has traditionally been used to estimate the locations of the support points and the masses at those points \citep{laird1978nonparametric}, but this is a difficult nonconvex optimization problem.

Instead, we maximize the pairwise composite likelihood over a fixed grid of support points, similar to recent $g$-modeling procedures for standard compound decision problems; this restores convexity \citep{jiang2009general, koenker2014convex, feng2018approximate}. Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$. We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
\[
\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
\]
over $k$. Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}. Our proposed estimator \ref{proposed} then becomes
\[
\hat{\bs{\delta}}(\bs{X})
=
(\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
\quad
\hat{t}(\bs{A}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}.
\]

Ideally, the grid points should be chosen to densely cover the parameter space. However, the fact that $G_p$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$. Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\hat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$. This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.

@@

\subsection{\label{posdef}Positive definiteness correction}
Our proposed estimator \ref{proposed} is not guaranteed to be positive-definite. To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix. \citet{higham1988computing} and \citet{huang2017calibration} showed that the projection of a $p \times p$ symmetric matrix $\bs{B}$ onto the space of positive semi-definite matrices is
\[
P_0(\bs{B})
=
\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert
=
\bs{Q}\text{diag}\{\max(\lambda_1,0),\max(\lambda_2,0),\ldots,\max(\lambda_p,0)\}\bs{Q}^\top,
\]
where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues. Projections in terms of other matrix norms are also possible.

@@

In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$. We generated $200$ replicates and reported average errors under the following three norms, where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
\begin{itemize}
\item Frobenius: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \ref{frobenius risk},

\item Spectral: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\hat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\hat{\bs{\Sigma}} - \bs{\Sigma}$, and

\item Matrix $\ell_1$: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \hat{\sigma}_{jk} - \sigma_{jk} \vert$.
\end{itemize}

@@

In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness. We compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:

@@

\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \ref{separable}. It equals our proposed estimator \ref{proposed} except with the true $G_p$ \ref{bayesian} instead of $\hat{G}_p$ \ref{Gp hat}. The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
\end{itemize}

\begin{figure}
\begin{center}
\centerline{  \includegraphics[width=0.85\textwidth]{img/sim2_frobenius.pdf}}
\end{center}
\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
\label{fig:sim2_frobenius}
\end{figure}

>>>>

\title[Empirical Bayes Covariance Matrix Estimation]{A Compound Decision Approach to Covariance Matrix Estimation}

@@

Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
\begin{equation}
\label{linear model}
\widehat{\bs{\Sigma}}_{\text{LW}} = \widehat{\rho}_{\text{LW}} \bs{S} + (1-\widehat{\rho}_{\text{LW}})\widehat{\mu} \bs{I}
\end{equation} 
where $\widehat{\mu} = \text{tr}(\bs{S})/p$, and $\widehat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\widehat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$. Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \ref{frobenius risk}. We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
\begin{equation}
  \label{linear class}
  \beta_S s_{jk} + \beta_I u_{jk},
\end{equation}
where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. It is straightforward to show that
$$\widehat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
is an unbiased estimate of the risk \ref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$. The optimal estimator in this class can now be chosen by minimizing $\widehat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}. Proposition \ref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \ref{linear model}.
\begin{prop}
  \label{prop:linear}
  Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{R}(\beta_S, \beta_I)$. Then $\widehat{\bs{\Sigma}}_V$ = $\widehat{\bs{\Sigma}}_{LW}$.
\end{prop}

@@

The previous section argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, but discusses only estimators linear in $s_{jk}$. We propose to consider a larger class, the class of so-called separable rules. In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ \citep{robbins1951asymptotically}. Here we generalize this to the problem of estimating a vectorized matrix. For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, define the class of separable rules
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}), j\neq k,\quad 
   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),j=1,\ldots,p)\},
\end{equation}
where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature. In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$. The sufficient statistics of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is $2\times 2$ matrix
$\begin{bmatrix}
s_{jj} & s_{jk} \\
s_{kj} & s_{kk} 
\end{bmatrix}$.

@@

The optimal $\bs{\delta}^\star$ is an oracle estimator and cannot be calculated in practice, as the true risk is unknown. In the classical compound decision framework, empirical Bayes methods are used to estimate the oracle optimal separable rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}. We take a similar approach here. To simplify notation, denote $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})^\top$ as $\bs{A}_{jk}$. $f(\cdot \mid \bs{\eta}_{jk})$, the density of $\bs{A}_{jk}$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation. When $r_{jk} \ne 1$, $\bs{A}_{jk}$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
\[
\bs{C}_{jk}=\begin{bmatrix}
&\sigma_j^2 & \sigma_j\sigma_k r_{jk} \\
& \sigma_j\sigma_k r_{jk}  & \sigma_k^2
\end{bmatrix}.
\]
When $r_{jk} = 1$, $\bs{A}_{jk}$ consists of mean-zero univariate normals with variances $\sigma_j^2$.

Now consider the Bayesian model for non-diagonal entries
\begin{equation}
  \label{bayesian nondiagonal}
  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k<j\leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
and diagonal entries
\begin{equation}
  \label{bayesian diagonal}
  \bs{A} \mid a \sim \widetilde{f}(\cdot \mid a),
  \quad
  a \sim G_d(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
\end{equation}
By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \ref{frobenius risk}: for any separable $\bs{\delta} \in \mathcal{S}$ \ref{separable}, the Frobenius risk can be written as
\begin{align*}
  R(\bs{\Sigma}, \bs{\delta})
  =\,&
  \frac{1}{p^2} \{\sum_{1\leq k<j\leq p}^p
  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A} 
  +
  \sum_{j=1}^p
  \int \{\widetilde{t}(\bs{X}) - \sigma_{jj} \}^2 \widetilde{f}(\bs{X} \mid s_{jj}) d\bs{X}
  \}\\
  =\,&
  \frac{p-1}{p}\int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_{nd}(\bs{\eta}) d\bs{A}
  +
  \frac{1}{p}\int \int \{\widetilde{t}(\bs{X}) - a^2\}^2 \widetilde{f}(\bs{X} \mid a) dG_{d}(a) d\bs{X} \\
  =\,&
   \frac{p-1}{p}\bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2]+ \frac{1}{p}\bb{E}[\{\widetilde{t}(\bs{X}) -  a^2\}^2],
\end{align*}
where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$. The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$ for $j\neq k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ and $\widetilde{t}^\star = \mathbb{E}\{g(a) \mid \bs{X}\}$ minimizes the Bayes risk.

Based on this result, we propose the following empirical Bayes procedure. We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the prior $G_p$. Under the Bayesian model \ref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_{nd}$ using
\begin{equation}
  \label{Gnd hat}
  \widehat{G}_{nd} = \argmax_{G \in \mathcal{G}_{nd}} \prod_{1\leq k<j<p} \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
\end{equation}

estimate $G_d$ using
\begin{equation}
  \label{Gd hat}
  \widehat{G}_d = \argmax_{G \in \mathcal{G}_d} \prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG(a),
\end{equation}
where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $\mathcal{G}_{d}$ is the family of all distributions supported on $\mathbb{R}$. Of course, the $\bs{A}_{jk}$ are not independent, so $\widehat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\widehat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using $ \widehat{\bs{\delta}}(\bs{X})
  =
  (\widehat{t}(\bs{A}_{21}), \ldots,  \widehat{t}(\bs{A}_{p,p-1}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$, where 
\begin{equation}
  \label{proposed}
  \widehat{t}(\bs{A}_{jk}) = \frac{\int g(\bs{\eta}) f(\bs{A}_{jk} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{A}_{jk} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
  \quad 
  \widehat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\widehat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\widehat{G}_d(a)}.
\end{equation}
The $\widehat{t}$ estimates the Bayes rule $t^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.

Our proposed procedure is an example of what \citet{efron2014two} calls $g$-modeling, an approach to empirical Bayes problems that proceeds by modeling the prior. A major advantage of nonparametric estiation of the prior is that it allows the data itself to determine how best to shrink the estimator. In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}. Theoretical justification of our proposed $\widehat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}. Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\widehat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.

@@

Calculating the estimated prior $\widehat{G}_p$ \ref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$. \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points. The EM algorithm has traditionally been used to estimate the locations of the support points and the masses at those points \citep{laird1978nonparametric}, but this is a difficult nonconvex optimization problem.

Instead, we maximize the pairwise composite likelihood over a fixed grid of support points, similar to recent $g$-modeling procedures for standard compound decision problems; this restores convexity \citep{jiang2009general, koenker2014convex, feng2018approximate}. Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$. We can then use the EM algorithm to estimate the masses $\widehat{\bs{w}}=\{\widehat{w}_{1},\ldots, \widehat{w}_{D}\}$ at those points via the iteration
\[
\widehat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\widehat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\widehat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
\]
over $k$. Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}. Our proposed estimator \ref{proposed} then becomes
\[
\widehat{t}(\bs{A}_{jk}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \widehat{w}_{\tau}},
\quad
\widehat{t}(\bs{X}_{\cdot j}) = \frac{ \sum_{\tau=1}^D a_{\tau}^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D \widetilde{f}(\bs{X}_{\cdot j}  \mid a_\tau) \widehat{w}_{\tau}}.
\]

Ideally, the grid points should be chosen to densely cover the parameter space. However, the fact that $G_p$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$. Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\widehat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$. This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.

@@

In our implementation, we use the density function of $\bs{S}_{jk}=\begin{bmatrix}
s_{jj} & s_{jk} \\
s_{kj} & s_{kk} 
\end{bmatrix}$ instead of $\bs{A}_{jk}$ since it is the sufficient statistics of $\bs{A}_{jk}$ and easier to calculate. Based on the property of sufficient statistics, 
\begin{equation}
\frac{f(\bs{A}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{A}_{jk}\mid \bs{\xi}_l)} = \frac{f(\bs{S}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{S}_{jk}\mid \bs{\xi}_l)}
\end{equation}
So it is equivalent to use $f(\bs{A}_{jk}\mid \bs{\xi}_{\tau})$ in calculation.
\subsection{\label{posdef}Positive definiteness correction}
Our proposed estimator \ref{proposed} is not guaranteed to be positive-definite. To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix. \citet{higham1988computing} and \citet{huang2017calibration} showed that the projection of a $p \times p$ symmetric matrix $\bs{B}$ onto the space of positive semi-definite matrices is
\[
P_0(\bs{B})
=
\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert
=
\bs{Q}\text{diag}\{\max(\lambda_1,0),\max(\lambda_2,0),\ldots,\max(\lambda_p,0)\}\bs{Q}^\top,
\]
where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues.

@@

In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$. We generated $200$ replicates and reported average errors under the following three norms, where $\widehat{\bs{\Sigma}}$ is the estimated matrix with entries $\widehat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
\begin{itemize}
\item Frobenius: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\widehat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \ref{frobenius risk},

\item Spectral: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\widehat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\widehat{\bs{\Sigma}} - \bs{\Sigma}$, and

\item Matrix $\ell_1$: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \widehat{\sigma}_{jk} - \sigma_{jk} \vert$.
\end{itemize}

@@

In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness. $K$-means clustering is applied on sample grids with $K=p$. We compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:

@@

\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \ref{separable}. It equals our proposed estimator \ref{proposed} except with the true $G_p$ \ref{bayesian} instead of $\widehat{G}_p$ \ref{Gp hat}. The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
\end{itemize}

\begin{figure}
\begin{center}
\centerline{  \includegraphics[width=0.85\textwidth]{img/sim2_frobenius.pdf}}
\end{center}
\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively. In MSG, $K$-means clustering is applied with $K=p$}
\label{fig:sim2_frobenius}
\end{figure}
